Prospective testing for the prevalence or transience of a shock effect over a set time horizon before it occurs

### 2022-07-30 prevalence_testing/postshocktest2022draft.tex
adjustments to employment series

The pandemic-induced recession of 2020 coincided with a one-month drop in US nonfarm payrolls in excess of 20 million people, or $13.57\%$, between March and April 2020.  This decline exceeds all other monthly declines by many orders of magnitude, and hence then distance-based weighting employed in \citep{lin2021minimizing} could not hope to recover the full shock effect.  However, the test developed herein does not rely as heavily upon the assumption of a common shock distribution for donors or the assumption that the donors represent a rich and diverse sampling from that distribution (\textcolor{red}{Think about this}).  Instead, what is needed is that the donor pool serve as an information source about the persistence of a shock effect in the time series under study, an information source based on the covariates that are posited to underlie the post-shock model dynamics (?).

We assemble a donor pool of substantial shocks to the US economy in the last four decades: the recessions of 1981-82, 2001-2003, and 2007-2009. These events bear similarity to the broad, seismic shock dealt to the US economy in March 2020 -- and the labor market in particular --  not only due to their unprecedented nature but also because of their potential to cause deep and lasting damage.  The recession of 1981-82 resulted from Fed Chairman Volcker's unprecedented and controversial bid to halt persistently high inflation that had reigned for more than decade in the US.  The economic shock that followed the September 11th terrorist both damaged consumer confidence and accelerated a contraction already in progress owing to the dot-com bubble.  Third and finally, the spring 2008 collapse of the investment bank Bear Stearns dealt a serious blow to confidence in the US financial system, which at the time was predicted by many to withstand the unprecedented declines in major US metro real estate markets.

We used several monthly macroeconomic indicators as covariates for the linear model as for arriving at an optimal weighting of the donor pool.  These indicators are all transformed to their month-on-month change in log values, so as to be consistent with the time series under study and to capture the signal inherent in changes, not raw levels, of macroeconomic variables.  These indicators are the US unemployment insurance transfers, real personal income, personal consumption expenditures, industrial production, consumer price index, Federal Reserve funds rate, and the count of black Americans age 20 and over on US nonfarm payrolls.

@@

In Table 3, using the synthetic pairwise model selection procedures specified in Corollary \ref{coro2}, because $\sum_{i=2}^4 w_i^* I(a_i^1 > a_i^2) = 0$,  we conclude that  $\mc{M}_{0}$ is preferred. Thus, we turn our interest to that model.  Since $p_{1}$ is far above the rejection threshold for horizons H=1,2,3,4, we lack the evidence to reject the null of a transient shock in the time series under study at any horizon.  The weighted p-value, $\hat p $, misses the ground truth p-value by no more than a modest amount at almost all horizons, and at some horizons is impressively close to the ground truth.  Last, it is worth remarking the for the preferred model $\mc{M}_{0}$, we observe a monotone increasing sequence of ground-truth p-values as H increases.  Such a phenomenon can be taken as evidence that as more data is harnessed, the decision is stable.  The estimated p-value $\hat p_{1}$ do an adequate job at tracking this phenomenon as H increases.

@@

\textcolor{red}{needs work} Talk about post-shock forecasting as parameterizing exogeniety and using hindsight in the donor pool to explicitly estimate exogeneity apriori. This framing has its limits, we are not able to perform post-shock forecasts when we are temporarily removed from the known shock time. For example, in the Conoco Philips analysis it is hard to imagine a successful post-shock forecast being made in the days prior to the weekend preceding Monday, March 9th 2020 as the OPEC-Russia oil supply increases and Covid-19 fears had not yet taken shape.

>>>>

The pandemic-induced recession of 2020 coincided with a one-month drop in US nonfarm payrolls in excess of 20 million people, or $13.57\%$, between March and April 2020.  This decline exceeds all other monthly declines by orders of magnitude, and hence then distance-based weighting employed in \citep{lin2021minimizing} could not hope to recover the full shock effect.  However, the test developed herein does not rely as heavily upon the assumption of a common shock distribution for donors or the assumption that the donors represent a rich and diverse sampling from that distribution (\textcolor{red}{Think about this}).  Instead, what is needed is that the donor pool serve as an information source about the persistence of a shock effect in the time series under study, an information source based on the covariates that are posited to underlie the post-shock model dynamics (?).

We assemble a donor pool of substantial shocks to the US economy in the last four decades: the recessions of 1981-82, 2001-2003, and 2007-2009. These events bear similarity to the broad, seismic shock dealt to the US economy in March 2020 -- and the labor market in particular --  not only due to their unprecedented nature but also because of their potential to cause deep and lasting damage.  The recession of 1981-82 resulted from Fed Chairman Volcker's unprecedented and controversial bid to halt persistently high inflation that had reigned for more than decade in the US \citep{sablik_2013}.  The economic shock that followed the September 11th terrorist both damaged consumer confidence and accelerated a contraction already in progress owing to the dot-com bubble \citep{kliesen_2003}.  Third and finally, the spring 2008 collapse of the investment bank Bear Stearns dealt a serious blow to confidence in the US financial system.  Prior to that, financial markets had withstood the unprecedented declines in major US metro real estate markets that followed the peak of the market in 2006.

We used several monthly macroeconomic indicators as covariates for the linear model to arrive at an optimal weighting of the donor pool.  These indicators are all transformed to their month-on-month change in log values, so as to be consistent with the time series under study and to capture the signal inherent in changes, not raw levels, of macroeconomic variables.  These indicators are the US unemployment insurance transfers, real personal income, personal consumption expenditures, industrial production, consumer price index, Federal Reserve funds rate, and the count of black Americans age 20 and over on US nonfarm payrolls.

@@

In Table 3, using the synthetic pairwise model selection procedures specified in Corollary \ref{coro2}, because $\sum_{i=2}^4 w_i^* I(a_i^1 > a_i^2) = 0$,  we conclude that  $\mc{M}_{0}$ is preferred. Thus, we turn our interest to that model.  Since $p_{1}$ is far above the rejection threshold for horizons H=1,2,3,4, we lack the evidence to reject the null of a transient shock in the time series under study at any horizon.  The weighted p-value, $\hat p $, misses the ground truth p-value by no more than a modest amount at almost all horizons, and at some horizons is impressively close to the ground truth.  Last, it is worth remarking the for the preferred model $\mc{M}_{0}$, we observe a monotone increasing sequence of ground-truth p-values as H increases.  Such a phenomenon can be taken as evidence that as more data is harnessed, the decision is stable.  The estimated p-value $\hat p_{1}$ do an adequate job at tracking this phenomenon as H increases.  The performance of method is in spite of the unusual path that the pre-lockdown US economy took, where weeks of restrictions on human assocation and movement gave way to relaxation of controls and an economic boom.

@@

\textcolor{red}{needs work} Talk about post-shock forecasting as parameterizing exogeneity and using hindsight in the donor pool to explicitly estimate exogeneity apriori. This framing has its limits, we are not able to perform post-shock forecasts when we are temporarily removed from the known shock time. For example, in the Conoco Philips analysis it is hard to imagine a successful post-shock forecast being made in the days prior to the weekend preceding Monday, March 9th 2020 as the OPEC-Russia oil supply increases and Covid-19 fears had not yet taken shape.
